---
title: "Working with Gene Objects in R using myGenePackage"
author: "Niccolò C.M. Bianchi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Working with Gene Objects in R using myGenePackage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r style, echo = FALSE, results = 'asis'}
library(BiocStyle)
```

```{r, echo = FALSE}
library(knitr)
library(GenomicRanges)
```

```{r, echo = FALSE}
library(geneObjects)
```


# Introduction

The **r Rpackage("myGenePackage")** package provides classes for representing different types of gene objects in R and associated methods. The package defines a parent class **Gene** and three subclasses **ProteinCodingGene**, **LncRNAGene**, and **MicroRNAGene**. It also provides methods to get and set the ID of the genes and to compute the length of gene products.


# Classes and Constructors

The following object is built using the *GenomicRanges* BioConductor package.

```{r}
gene_structure <- GenomicRanges::GRanges(seqnames = "chr1", ranges = IRanges::IRanges(start = 5000, end = 6000), strand = "+")
```

## Protein-coding genes

The **ProteinCodingGene** S4 class represents protein-coding genes. It can be created using the *ProteinCodingGene* constructor:

```{r}
# create a ProteinCodingGene object
pc_gene <- ProteinCodingGene(ID = "ENSG000001", HUGO_symbol = "BRCA1", gene_name = "Breast cancer 1", description = "Breast cancer type 1 susceptibility protein", gene_structure = gene_structure, protein_ID = "ENSP000003", protein_sequence = "MENSDRNSIKVAV")
```

## Long non-coding RNA gene

The **LncRNAGene** class represents long non-coding RNA genes. It can be created using the *LncRNAGene* constructor:

```{r}
# create a LncRNAGene object
lnc_gene <- LncRNAGene(ID = "ENSG000002", HUGO_symbol = "LINC00001", gene_name = "Long intergenic non-protein coding RNA 1", description = "Long non-coding RNA", gene_structure = gene_structure, lncRNA_ID = "ENST000004", RNA_sequence = "ACUGCUAGCUAGUCA")
```

## MicroRNA gene

The **MicroRNAGene** class represents microRNA genes. It can be created using the *MicroRNAGene* constructor:

```{r}
# create a MicroRNAGene object
mir_gene <-  MicroRNAGene(ID = "ENSG000003", HUGO_symbol = "MIR00001", gene_name = "MicroRNA 1", description = "MicroRNA", gene_structure = gene_structure, microRNA_ID = "ENST000005", microRNA_seed_sequence = "UGAGGUAGUAGGUUGUAUGGUAG")
```


# Gene ID getter and setter

The package provides generic functions to get and set the ID of a gene. Here is an example of how to use these functions:

```{r}
# set ID
setID(pc_gene) <- "Gene1"

# get ID
getID(pc_gene)
```


# Length of gene product

The **lengthProduct** function computes the length of the product of a gene. Here is an example of how to use this function:

```{r}
# compute length of gene product
lengthProduct(pc_gene)
```
